Multi-resolution Neural Field Estimation

### 2011-03-07 Multiresolution_framework/ltx/manuscript.tex
have compressed a bit the final two paragraphs of the conclusion.

\parham{The actual connectivity kernel} and the decomposition is plotted in \figurename{\ref{fig:KernelEstimate}}(a). The reconstructed kernel is in good accordance with the actual kernel, where the actual kernel lies inside the confidence interval. The small error in the estimate is attributed to the MRA of the system used to form the estimator. \figurename{\ref{fig:KernelEstimate}}(b) and (c) shows the kernel scaling and wavelet functions, respectively.
\begin{figure}[!h] 
	\centering
		\includegraphics[scale=1]{./Graph/Field.pdf}
		\caption{Actual (grey) and estimated (black) neural fields at a time instant for different spatial resolutions. (a) $j=0$, $n_x=17$, $RMSE = 2.10$. (b) $j=1$, $n_x=33$, $RMSE = 1.42$. (c) $j=2$, $n_x=65$, $RMSE = 0.80$. (d) $j=3$, $n_x=131$, $RMSE = 0.72$. (e) $j=4$, $n_x=263$, $RMSE=0.72$.}
	\label{fig:FieldEstimates}
\end{figure} 
\begin{figure}[!h] 
	\centering
		\includegraphics[scale=1]{./Graph/KernelUnknownSupportWithVarsigma.pdf}
		\caption{Connectivity kernel estimate. (a) The actual kernel is shown with the black solid line. The estimated kernel and confidence interval is shown by the red line and red shaded region ($\pm2$ std.). (b) Kernel scaling functions at $j=1$. (c) Kernel wavelet functions at $j=1$.}
	\label{fig:KernelEstimate}
\end{figure}
% \caption{The actual connectivity kernel is shown with the black solid line. The estimated kernel and confidence interval is shown by the red line and red shaded region ($\pm2$ std.). The 7 weighted kernel basis functions are shown by the black dashed lines.}
\section{Discussion}
In this paper, we have presented a novel model-based framework for estimating cortical dynamics from electrophysiological measurements. The novel and key developments of the paper include the multi-resolution FEM representation of the neural field, and the estimator for an intracortical connectivity kernel with an arbitrary shape. This work is significant, as the ability to create patient specific neural field models has the potential to contribute to our understanding and improve treatment of diseases resulting from abnormal neurodynamics, such as epilepsy. Other groups have previously highlighted the importance of a multi-resolution approach in neural field modeling. For example, it is thought that the dynamics and connectivity structure differs at different spatial scales~\cite{Qubbaj2009}.

@@

In order to apply the framework to real data some assumptions must be made, such as stationarity of the parameters over the estimation period. Perhaps the most critical assumption is that the model provides an apt description of the cortical dynamics. The authors acknowledge that there is, and will always be, a discrepancy between the model and cortex. Nevertheless, the model-based framework proposed in this paper may enable meaningful state tracking and connectivity estimation. The key development in this paper is the multi-resolution decomposition forming the state-space model. The decomposition holds for more sophisticated nonlinear forms of the field equations, where the unscented RTS smoother should be used in the E-step of the estimation algorithm.

Future work should be directed towards extending the framework to account for a second-order synaptic response kernel, heterogeneity in the connectivity and perhaps time delays at lower spatial resolutions. Most importantly, future work should be directed towards applying and validating the framework on real data.

>>>>

\parham{The actual connectivity kernel} and the decomposition is plotted in \figurename{\ref{fig:KernelEstimate}}(a). The reconstructed kernel is in good accordance with the actual kernel, where the actual kernel lies inside the confidence interval. The small error in the estimate is attributed to the MRA of the system used to form the estimator. \figurename{\ref{fig:KernelEstimate}}(b) and (c) shows the kernel scaling and wavelet functions, respectively.
\begin{figure}[!h] 
	\centering
		\includegraphics[scale=1]{./Graph/Field.pdf}
		\caption{Actual (grey) and estimated (black) neural fields at a time instant for different spatial resolutions. (a) $j=0$, $n_x=17$, $RMSE = 2.10$. (b) $j=1$, $n_x=33$, $RMSE = 1.42$. (c) $j=2$, $n_x=65$, $RMSE = 0.80$. (d) $j=3$, $n_x=131$, $RMSE = 0.72$. (e) $j=4$, $n_x=263$, $RMSE=0.72$.}
	\label{fig:FieldEstimates}
\end{figure} 
\begin{figure}[!h] 
	\centering
% 		\includegraphics[scale=1]{./Graph/KernelUnknownSupportWithVarsigma.pdf}
		\includegraphics[scale=1]{./Graph/test.pdf}
		\caption{Connectivity kernel estimate. (a) The actual kernel is shown with the black solid line. The estimated kernel and confidence interval is shown by the red line and red shaded region ($\pm2$ std.). (b) Kernel scaling functions at $j=1$. (c) Kernel wavelet functions at $j=1$.}
	\label{fig:KernelEstimate}
\end{figure}
% \caption{The actual connectivity kernel is shown with the black solid line. The estimated kernel and confidence interval is shown by the red line and red shaded region ($\pm2$ std.). The 7 weighted kernel basis functions are shown by the black dashed lines.}
\section{Discussion}
In this paper, we have presented a novel model-based framework for estimating cortical dynamics from electrophysiological measurements. The novel and key developments of the paper include the multi-resolution FEM representation of the neural field, and the estimator for an intracortical connectivity kernel with an arbitrary shape. This work is significant, as the ability to create patient specific neural field models has the potential to contribute to our understanding and improve treatment of diseases resulting from abnormal neurodynamics, such as epilepsy. Other groups have previously highlighted the importance of a multi-resolution approach in neural field modeling. For example, it is thought that the dynamics and connectivity structure differs at different spatial scales~\cite{Qubbaj2009}.

@@

\mike{The presented framework makes two key assumptions about the cortex: a linear activiation function and stationary dynamics. Our main aim for this framework is to relax these assumptions using the multi-resolution decomposition. While the decomposition itself holds, efficiently performing nonlinear smoothing in the much larger state space remains a challenge. Additional avenues for future work include extending the model to capture a second order synaptic response kernel, and time delays at lower spatial resolutions.  The estimation techniques should also be extended to deal with spatial and temporal heterogeneity in the kernel. Finally, the majority of our current work involves applying this framework to collected data, primarily for the use of seizure detection and mitigation.}

\cut{In order to apply the framework to real data some assumptions must be made, such as stationarity of the parameters over the estimation period. Perhaps the most critical assumption is that the model provides an apt description of the cortical dynamics. The authors acknowledge that there is, and will always be, a discrepancy between the model and cortex. Nevertheless, the model-based framework proposed in this paper may enable meaningful state tracking and connectivity estimation. The key development in this paper is the multi-resolution decomposition forming the state-space model. The decomposition holds for more sophisticated nonlinear forms of the field equations, where the unscented RTS smoother should be used in the E-step of the estimation algorithm.

Future work should be directed towards extending the framework to account for a second-order synaptic response kernel, heterogeneity in the connectivity and perhaps time delays at lower spatial resolutions. Most importantly, future work should be directed towards applying and validating the framework on real data.}
